---
title: "Individual Analysis"
author: "Samuel Reinart"
date: "2/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Searching for the Best Chocolates

```{r}
library(dplyr)

choco <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/data/choco.csv")

ratings <- choco$Rating
companies <- choco$Company
locationDF <- count(choco, Location)
locationCount <- setNames(as.integer(locationDF$n), substr(locationDF$Location, 1, 13))
cacaoRatings <- choco[c("Cocoa_Percent", "Rating")]
head(choco)
```

1.
```{r}
hist(ratings)
boxplot(ratings)
```

As you can see from the histogram, most of the chocolate bars have a rating between 2.75 and 3.75. From the boxplot shows that the median rating is around 3.25. There are a few outliers with low ratings on the boxplot. The histogram tells us that there are around 50 chocolate bars with ratings of 2 or less.

2. 
```{r}
locationDF <- count(choco, Location)
locationCount <- setNames(as.integer(locationDF$n), substr(locationDF$Location, 1, 13))
barplot(locationCount, las=2, cex.names=0.8, ylim=c(0,1000))
barplot(locationCount, las=2, cex.names=0.3, ylim=c(0,1000))
```

A large majority of the ratings were made for companies located in the United States. This is likely due to the fact that there are a lot of chocolate companies based in the U.S. The next closest countries are France and Canada. The rest of the countries have very few ratings, likely because they have much fewer chocolate companies.

3. 
```{r}
plot(cacaoRatings)
cacaoMeans <- aggregate(Rating ~ Cocoa_Percent, cacaoRatings, mean)
cacaoMeansV <- setNames(cacaoMeans$Rating, cacaoMeans$Cocoa_Percent)
barplot(cacaoMeansV, las=2)
cor(cacaoMeans$Rating, cacaoMeans$Cocoa_Percent)
```

A scatter plot is not a good choice for visualization because it is difficult to see a pattern or trend on the graph. However, after taking the average score of each cacao percentage and plotting it using a boxplot, we can see that there does not appear to be a direct relationship between the rating of chocolate and its cocoa percentage. We can confirm this by seeing the correlation between the two values. Since the correlation value is low, we can determine that the cocoa percentage does not have a large impact on the rating a chocolate bar receives.

4.
```{r}
originRatings <- choco[c("Bean_Origin", "Rating")]
originMeans <- aggregate(Rating ~ Bean_Origin, originRatings, mean)
originMeansV <- setNames(originMeans$Rating, originMeans$Bean_Origin)
barplot(originMeansV, las=2, cex.names=0.8)
head(originMeans[order(originMeans$Rating), ])
head(originMeans[order(-originMeans$Rating), ])
```

There doesn't seem to be a lot of variation in the ratings of beans, as seen in the barplot. Many of the countries from which the highest and lowest rated beans were in the Caribbean and near Africa according to the sorted data frame. However, two countries that produce some of the highest rated beans, Thailand and the Solomon Islands, break this trend since they are located in Asia and Oceania.

5. 
```{r}
numIngredientsRatings <- choco[c("Ingredients", "Rating")]
numIngredientsRatings$Num_Ingredients <- with(numIngredientsRatings, as.numeric(substr(Ingredients, 1, 1)))
numIngredientsRatings <- numIngredientsRatings[,c("Num_Ingredients", "Rating")]
numMeans <- aggregate(Rating ~ Num_Ingredients, numIngredientsRatings, mean)
numMeansV <- setNames(numMeans$Rating, numMeans$Num_Ingredients)
barplot(numMeansV, las=2, cex.names=0.8)
cor(numMeans$Rating, numMeans$Num_Ingredients, method="spearman")
```

I decided to analyze the data to see if the number of ingredients has an effect on the rating a chocolate bar receives. The bar plot shows us that chocolate containing 2 or 3 ingredients has the highest average rating, with 3 being the best. More or less ingredients negatively impacts the score. The correlation between the rating and the number of ingredients seems to support this, as the correlation was -0.37 which suggests a moderate correlation between the two variables.

```{r}
allMean <- mean(choco$Rating)
vanilly <- choco[c("Ingredients", "Rating")]
vanillaRatings <- filter(vanilly, grepl("V", Ingredients))
nonVanillaRatings <- filter(vanilly, !grepl("V", Ingredients))
vanillaMean <- mean(vanillaRatings$Rating)
nonVanillaMean <- mean(nonVanillaRatings$Rating)
vanillaTest <- c(allMean, vanillaMean, nonVanillaMean)
names(vanillaTest) <- c("All Chocolate", "Chocolate With Vanilla", "Chocolate Without Vanilla")
vanillaTest
barplot(vanillaTest, ylim=c(0,3.5))
```

I also analyzed the affect of vanilla on the rating of chocolate. I found that the inclusion of vanilla as one of the ingredients of chocolate negatively impacts the taste of the chocolate. The average rating of chocolate with vanilla is 3.03, while the average rating of chocolate without vanilla is 3.21. This means that, on average, adding vanilla drops the rating of chocolate by 0.18.
